---
title: "Park_Lab4"
author: "Patricia Park"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(skimr)
library(tidymodels)
library(caret)
library(corrplot)
```

## Lab 4: Fire and Tree Mortality

The database we'll be working with today includes 36066 observations of individual trees involved in prescribed fires and wildfires occurring over 35 years, from 1981 to 2016. It is a subset of a larger fire and tree mortality database from the US Forest Service (see data description for the full database here: [link](https://www.nature.com/articles/s41597-020-0522-7#Sec10)). Our goal today is to predict the likelihood of tree mortality after a fire.

### Data Exploration

Outcome variable: *yr1status* = tree status (0=alive, 1=dead) assessed one year post-fire.

Predictors: *YrFireName, Species, Genus_species, DBH_cm, CVS_percent, BCHM_m, BTL* (Information on these variables available in the database metadata ([link](https://www.fs.usda.gov/rds/archive/products/RDS-2020-0001-2/_metadata_RDS-2020-0001-2.html))).

```{r}
trees_dat<- read_csv(file = "https://raw.githubusercontent.com/MaRo406/eds-232-machine-learning/main/data/trees-dat.csv")
```

> Question 1: Recode all the predictors to a zero_based integer form

```{r}
#specify and prep recipe
trees_rec <- recipe(yr1status ~., data = trees_dat) %>%
  step_integer(all_predictors(), zero_based = T) %>%
  prep(trees_dat) %>% 
  bake(trees_dat)

```


### Data Splitting

> Question 2: Create trees_training (70%) and trees_test (30%) splits for the modeling

```{r}
set.seed(123)  # for reproducibility 
trees_split <- initial_split(trees_rec, prop = .7)
trees_train <- training(trees_split)
trees_test  <- testing(trees_split)
```


> Question 3: How many observations are we using for training with this split?

```{r}
tree_train_ob <- nrow(trees_train)
print(tree_train_ob)
```

**Answer**: We have `r tree_train_ob` observations that we are using for the training data.


### Simple Logistic Regression 

Let's start our modeling effort with some simple models: one predictor and one outcome each.

> Question 4: Choose the three predictors that most highly correlate with our outcome variable for further investigation.

```{r}
# Obtain correlation matrix
corr_mat <- cor(trees_rec %>% 
                  # Drop columns that are not really informative
                  select(-c(`...1`)))

# Make a correlation plot between the variables
corrplot(corr_mat, method = "shade", shade.col = NA, tl.col = "black", tl.srt = 45, addCoef.col = "black", cl.pos = "n", order = "original")

```


**Answer**: The three predictors that most highly correlate with our outcome variables `yr1status` are `CVS_percent`, `BCHM_m`, and `DBH_cm`.

> Question 5: Use glm() to fit three simple logistic regression models, one for each of the predictors you identified.

```{r}
#CVS_percent
model_CVS <- glm(data = trees_rec, yr1status ~ CVS_percent, family = "binomial")
  
#BCHM_m
model_BCHM <- glm(data = trees_rec, yr1status ~ BCHM_m, family = "binomial")

#DBH_cm
model_DBH <- glm(data = trees_rec, yr1status ~ DBH_cm, family = "binomial")

```

### Interpret the Coefficients 

We aren't always interested in or able to interpret the model coefficients in a machine learning task. Often predictive accuracy is all we care about.

> Question 6: That said, take a stab at interpreting our model coefficients now.

```{r}
broom::tidy(model_CVS)
broom::tidy(model_BCHM)
broom::tidy(model_DBH)
(coef(model_CVS))
(coef(model_BCHM))
(coef(model_DBH))
```

**Answer**: Looking at all three predictors, we know that they are all significant in impacting the `yr1status` because of how small the p-value is for all three of them. If we want to interpret the coefficients, every time the `CVS_percent` goes up by one unit, the `yr1status` goes up by 0.0774. Every time the `BCHM_m` goes up by one unit, the `yr1status` goes up by 0.00625. Every time the `DBH_cm` goes up by one unit, the `yr1status` goes down by -0.00370.


> Question 7: Now let's visualize the results from these models. Plot the fit to the training data of each model.

```{r}
ggplot(trees_train, aes(x = CVS_percent, y = yr1status)) + geom_point() +
  stat_smooth(method = "glm", se = T, method.args = list(family = binomial)) +
  labs(title = "Logistic Regression between pre-fire crown volume and tree status one year post-fire",
       x = "Percent of pre-fire crown volumn",
       y = "Tree status one year post-fire")

ggplot(trees_train, aes(x = BCHM_m, y = yr1status)) + geom_point() +
  stat_smooth(method = "glm", se = T, method.args = list(family = binomial)) +
  labs(title = "Logistic Regression between Trunk Diameter and tree status one year post-fire",
       x = "Trunk Diameter (cm)",
       y = "Tree status one year post-fire")

ggplot(trees_train, aes(x = DBH_cm, y = yr1status)) + geom_point() +
  stat_smooth(method = "glm", se = T, method.args = list(family = binomial)) +
  labs(title = "Logistic Regression between Maximum bark char and tree status one year post-fire",
       x = "Maximum bark char (m)",
       y = "Tree status one year post-fire")
```

*Note*: It's interesting to see that for the lm line for the BCHM_m, it follows a near perfect logistic line. The regression line for CVS_percent leans more towards zero as well as the DBH_cm graph.


### Multiple Logistic Regression (chapter 5)

Let's not limit ourselves to a single-predictor model. More predictors might lead to better model performance.

> Question 8: Use glm() to fit a multiple logistic regression called "logistic_full", with all three of the predictors included. Which of these are significant in the resulting model?

```{r}
logistic_full <- glm(data = trees_rec, yr1status ~ DBH_cm + CVS_percent + BCHM_m, family = "binomial")
broom::tidy(logistic_full)
```

**Answer**: While all of these predictors are significant in the resulting model, CVS_percent is more significant than the other predictors as the p_value is a near 0 as seen on the table. The estimate coefficient is also very high compared to the other predictors coefficients.


### Estimate Model Accuracy

Now we want to estimate our model's generalizability using resampling.

> Question 9: Use cross validation to assess model accuracy. Use caret::train() to fit four 10-fold cross-validated models (cv_model1, cv_model2, cv_model3, cv_model4) that correspond to each of the four models we've fit so far: three simple logistic regression models corresponding to each of the three key predictors (CVS_percent, DBH_cm, BCHM_m) and a multiple logistic regression model that combines all three predictors.

```{r, warning=FALSE}
#change the outcomes to factors
trees_train$yr1status <- as.factor(trees_train$yr1status)

#create the cross-validation modes for all predictors
#DBH_cm
cv_model1 <- train(yr1status~as.factor(DBH_cm),
                   data = trees_train,
                   method = 'glm',
                   family = 'binomial',
                   trControl = trainControl(method = 'cv', number = 10))

#CVS_percent
cv_model2 <- train(yr1status~ CVS_percent,
                   data = trees_train,
                   method = 'glm',
                   family = 'binomial',
                   trControl = trainControl(method = 'cv', number = 10))

#BCHM_m
cv_model3 <- train(yr1status ~BCHM_m,
                   data = trees_train,
                   method = 'glm',
                   family = 'binomial',
                   trControl = trainControl(method = 'cv', number = 10))

#DBH_cm, CVS_percent, and BCHM_m
cv_model4 <- train(yr1status ~DBH_cm + CVS_percent + BCHM_m,
                   data = trees_train,
                   method = 'glm',
                   family = 'binomial',
                   trControl = trainControl(method = 'cv', number = 10))

```

> Question 10: Use caret::resamples() to extract then compare the classification accuracy for each model. (Hint: resamples() wont give you what you need unless you convert the outcome variable to factor form). Which model has the highest accuracy?

```{r}
summary(resamples(list(model1 = cv_model1,
                       model2 = cv_model2,
                       model3 = cv_model3,
                       model4 = cv_model4)))
```

**Answer**: best model is model 4

Let's move forward with this single most accurate model.

> Question 11: Compute the confusion matrix and overall fraction of correct predictions by the model.

```{r}
pred_class <- predict(cv_model4, trees_train)
# making confusion matrix
confusionMatrix(
  data = relevel(pred_class, ref = '1'),
  reference = relevel(trees_train$yr1status, ref = "1"))

```


> Question 12: Explain what the confusion matrix is telling you about the types of mistakes made by logistic regression.
>
> Question 13: What is the overall accuracy of the model? How is this calculated?

### Test Final Model

Alright, now we'll take our most accurate model and make predictions on some unseen data (the test data).

> Question 14: Now that we have identified our best model, evaluate it by running a prediction on the test data, trees_test.
>
> Question 15: How does the accuracy of this final model on the test data compare to its cross validation accuracy? Do you find this to be surprising? Why or why not?
